\chapter*{Declaration}

I certify that the thesis presented here for examination for a Ph.D.\ degree of the University of Glasgow is my own work, with the following exceptions:
\begin{itemize}
    \item Chapter~\ref{c:three_dsm_cyc} includes joint work with Klaus Heeger. In particular, it was Klaus' idea that led to a significant improvement of the approximation ratio (in Theorem~\ref{thm:three_dsm_cyc_unrestricted}) and a collaborative brainstorm that produced the approximation algorithm in the setting of master lists (shown in Theorem~\ref{thm:three_dsm_cyc_ml}).
    \item The instance described in Theorem~\ref{thm:threed_sr_b_ifnis15thensmdoesnotexist}, presented in Chapter~\ref{c:threed_sr_b}, is the product of a joint effort with Sofia Simola. 
    \item Chapter~\ref{c:threed_efr_as} includes joint work with \'Agnes Cseh, conducted part-time from December 2020 to May 2022. \'Agnes provided expert guidance and advice throughout the project. In terms of our new results, it was her ideas that led to the polynomial-time algorithm and $\NP$-completeness reduction related to envy (now shown in Theorem~\ref{thm:threed_efr_as_regularenvy_npcomplete}). We worked jointly on the algorithm and reduction for weakly justified envy, which I then formalised and wrote up (in Theorem~\ref{thm:threed_efr_as_wjef_npcomplete}). I developed the $\NP$-completeness reduction for weakly justified envy (shown in Theorem~\ref{thm:threed_efr_as_wjef_algowjpathscycles}) and proved the results relating to justified envy (shown in Theorems~\ref{thm:threed_efr_as_jef_binary_algorithm}--\ref{thm:threed_efr_as_jef_symmetric_6_npcomplete}).  Everything else was a combined effort.
\end{itemize}

My supervisor, David Manlove, has supported this research throughout, providing general guidance and critically reviewing technical material. 

\chapter*{Research output}

The following papers are a result of the research presented in this thesis:
\begin{itemize}
    \item M.\ McKay and D.\ Manlove. The Three-Dimensional Stable Roommates Problem with Additively Separable Preferences. In \emph{Proceedings of SAGT '21: The 14\textsuperscript{th} International Symposium on Algorithmic Game Theory}. LNCS, vol.\ 12885, pp.\ 266--280. Springer (2021). % A full version is available as Technical Report number 2107.04368, arXiv (2021) from \url{https://arxiv.org/abs/2107.04368}.
    This paper contains some of the results presented in Chapter~\ref{c:threed_sr_as}.
    
    \item \'A. Cseh, M.\ McKay, and D.\ Manlove. Envy-freeness in 3D Hedonic Games. Submitted to \emph{Autonomous Agents and Multi-Agent Systems}, 2023. A preprint is available as Technical Report number 2209.07440 (2022) on arXiv, from \url{https://arxiv.org/abs/2209.07440}. This paper contains the results presented in Chapter~\ref{c:threed_efr_as}.
    
    \item M.\ McKay and D.\ Manlove. Packing $K_r$s in bounded degree graphs. Submitted to \emph{Discrete Applied Mathematics}, 2022. A preprint is available as Technical Report number 2209.03684 (2022) on arXiv, from \url{https://arxiv.org/abs/2209.03684}. This paper contains the results presented in Chapter~\ref{c:kr_packing}.
\end{itemize}