The remainder of this thesis is structured as follows:

\begin{itemize}
    \item In Chapter~\ref{c:lit_review}, we give an overview of literature involving coalition formation, hedonic games, and fixed-size coalitions, setting the scene for our subsequent results.
    
    \item In Chapter~\ref{c:three_dsm_cyc}, we introduce 3-DSM-CYC and present new results on the approximability of a related optimisation problem, both in the general case and in a restriction of the problem involving master lists.
    
    \item In Chapter~\ref{c:threed_sr_b}, we define 3DR-B and present new results concerning the existence of stable matchings in instances of 3DR-B and on the approximability of a related optimisation problem.
    
    \item In Chapter~\ref{c:threed_sr_w}, we define 3DR-W and present new results concerning the existence of stable matchings in instances of 3DR-W and on the approximability of a related optimisation problem.
    
    \item In Chapter~\ref{c:threed_sr_as}, we define 3DR-AS and present new results concerning the existence of stable matchings in 3DR-AS, under various restrictions on the agents' valuations.
    
    \item In Chapter~\ref{c:threed_efr_as}, we present new results concerning the existence of envy-free, weakly justified envy-free, and justified envy-free matchings in instances of 3DR-AS, under various restrictions on the agents' valuations.
    
    \item In Chapter~\ref{c:kr_packing}, we present new results on the complexity of the $K_r$-packing problem in graphs of fixed maximum degree.
    
    \item In Chapter~\ref{c:conclusion}, we summarise the contribution of this thesis and discuss some directions for future work.
\end{itemize}