In this chapter we formalised a model of 3DR involving additively separable preferences, which we called 3DR-AS. We considered in instances of 3DR-AS the existence of, and complexity of finding, matchings that are stable, under three possible restrictions of the agents' valuations. 

We first showed that any instance of 3DR-AS with binary and symmetric preferences contains a stable matching, and presented a polynomial-time algorithm that can construct a stable matching in such an instance. 
We then considered the problem of finding a stable matching with maximum utilitarian welfare, given an instance with binary and symmetric preferences. We called this problem the \emph{3DR-AS Stable Maximum Utilitarian Welfare problem} (\mysymbolfirstusedefinition{symboldef:threedr_as_smuw}{3DR-AS-SMUW}). We proved that 3DR-AS-SMUW is $\NP$-hard, but showed that the algorithm for constructing a stable matching in this setting can be modified to yield a $2$-approximation algorithm.  We then complemented the previous tractability results with two hardness results. The first is that a stable matching need not exist in general, and the associated decision problem is $\NP$-complete even when preferences are binary and not necessarily symmetric. The second is that the same decision problem is $\NP$-complete even when preferences are ternary and symmetric. 

Later, in Chapter~\ref{c:threed_efr_as}, we shall combine some of the results from this chapter with other results for 3DR-AS and present a comprehensive analysis of the existence and complexity of feasible matchings in 3DR-AS under restricted preferences, for four related solution concepts. This classification is shown in Table~\ref{tab:introduction_3dsras_mainresults} in Chapter~\ref{c:threed_efr_as}.

We now present some open problems specifically involving stability in 3DR-AS. More general problems, involving solution concepts other than stability and other models of fixed-size coalitions, are discussed in Chapter~\ref{c:conclusion}. 

An immediate open problem is whether our results for binary preferences hold in a more general setting in which $v_{\alpha_i}(\alpha_j) \in \{ a, b \}$ for any non-negative integers $a$ and $b$ where $a < b$, and whether our results for ternary preferences hold when $v_{\alpha_i}(\alpha_j) \in \{ a, b, c \}$ for any non-negative integers $a$, $b$, and $c$ where $a < b < c$. We conjecture that both statements are true, and that both the polynomial-time algorithm and $\NP$-hardness reductions can be modified accordingly.

As we noted in Section~\ref{sec:threed_sr_as_symmetric_binary_welfare}, any approximation algorithm for 3DR-AS-SMUW with approximation ratio strictly less than $2$ must not always begin, like Algorithm~\algorithmfont{findStableUW} does, by selecting a maximal set of triangles. While it would be very interesting to derive such an algorithm, it would also be informative to derive an inapproximability result for 3DR-AS-SMUW. We conjecture that 3DR-AS-SMUW is $\APX$-hard.

An open direction of work involves the \emph{price of anarchy} and \emph{price of stability} of 3DR-AS \cite{Bilo22}. In the setting of 3DR-AS, we could define the price of anarchy (stability) as the worst- (best-)case ratio between the utilitarian welfare of an arbitrary stable matching and the maximum possible utilitarian welfare over all matchings. Our example instance, shown in Figure~\ref{fig:threed_sr_as_max_welfare_example}, thus shows that the price of anarchy is at least $2$, even when preferences are binary and symmetric. 

It would be very interesting to identify other restrictions of 3DR-AS in which a stable matching can be found in polynomial time. For example, a crucial part of the reduction shown in Section~\ref{sec:threed_sr_as_generalbinary} was the pentagadget, which corresponds to a unique type of regular tournament graph \cite{McDonald2020}. It would be quite remarkable if, for example, every instance of 3DR-AS that does not contain a pentagadget contains a stable matching. As a starting point, we conjecture that any instance corresponding to a directed acyclic graph must contain a stable matching.

Continuing the connection to graph theory (discussed in Section~\ref{sec:threed_sr_as_intro}), it might also be possible to consider the problem of finding a stable matching in a given instance of 3DR-AS from the perspective of the parameterised complexity with respect to a graph parameter. For example, in the case of binary and symmetric preferences, one could consider the tree-width \cite{Robertson84} of the underlying graph.

As we noted for 3DR-B and 3DR-W (in Chapters~\ref{c:threed_sr_b} and~\ref{c:threed_sr_w}), it would be very interesting to estimate the probability that a random instance of 3DR-AS contains a stable matching, or to estimate the same probability in a random instance of 3DR-AS with binary or ternary preferences. It might be possible to use probabilistic techniques from graph theory, such as the Erd\H{o}s-R\'enyi model of a random graph. Alternatively an empirical approach might be informative, for example by formulating the problem as an integer program \cite{IPbook}.

% The connection between 3DR-AS and graph theory gives rise to other interesting research questions. For example, one could study the problem of finding a stable matching in a given instance of 3DR-AS with from the view of its parameterised complexity. A variety of parameters are possible, in particular in the setting of binary preferences, in which we can view an instance of 3DR-AS as a directed graph. For example, the maximum in- or out-degree. 
