In this chapter we formalised a new model of 3DR, involving $\mathscr{W}$-preferences, which we called 3DR-W. We considered in 3DR-W the existence of, and complexity of finding, matchings that are stable.

We first showed that, as in the case of 3DR-B (see Chapter~\ref{c:threed_sr_b}), a given instance of 3DR-W may not contain a stable matching and the associated decision problem is $\NP$-complete. Next, we considered a closely related optimisation problem, which we called 3DR-W-MSM, in which the objective is to construct, in a given instance of 3DR-B, a matching with the maximum number of non-blocking triples. Finally, we presented a $9/4$-approximation algorithm for 3DR-W-MSM built on an existing algorithm for 3PSA-MSM, which is a closely related problem \cite{rosenbaum16}. 

We now present some open problems specifically involving stability in 3DR-W. More general problems, involving solution concepts other than stability and other models of fixed-size coalition formation, are discussed in Chapter~\ref{c:conclusion}. 

% Although the definitions of 3DR-W and 3DR-B are similar, and it is $\NP$-complete to decide if a given instance of either model contains a stable matching, it appears that their commonality is otherwise limited. For example, it is straightforward to construct an instance of 3DR-W with $6$ agents that contains no stable matching (see Section~\ref{sec:threed_sr_w_hardness}) but, as we saw in Chapter~\ref{c:threed_sr_b}, the smallest such instance of 3DR-B has at at least $9$ (and at most $63$) agents. 

Although the definitions of 3DR-B and 3DR-W are similar, and the stability existence problem is $\NP$-complete in both models, it is unclear whether the two models are further related. For example, in Chapter~\ref{c:threed_sr_b} we presented a $3/2$-approximation algorithm for 3DR-B-MSM, but it seems difficult to design such an algorithm for 3DR-W-MSM. One possible starting point is Rosenbaum's $9/4$-approximation algorithm for 3PSA-MSM, Algorithm~\algorithmfont{ASA}. Another idea is to adapt Irving's \cite{Irv85} algorithm for (two-dimensional) Stable Roommates. In fact, Irving's algorithm is the basis of another polynomial-time algorithm for an analogous model involving  $\mathscr{W}$-preferences \cite{CH04}, in which coalitions need not have a fixed size.

Alternatively, it might be possible to prove an inapproximability result for 3DR-W-MSM. For example, if one could show that no $3/2$-approximation algorithm exists for 3DR-W-MSM, unless $\P=\NP$, then a striking difference would be revealed between 3DR-B and 3DR-W. To this end, it might be useful to consider the reductions of Iwama et al.\ \cite{IMO08} and Rosenbaum \cite{rosenbaum16}, which show that two optimisation problems that are defined in related models of 3DR are both $\APX$-hard (both results are discussed in Chapter~\ref{c:lit_review}).



It would be very informative to estimate the probability that a random instance of 3DR-W contains a stable matching. In particular, if this probability could be compared between the models of 3DR-B and 3DR-W. As we noted in Chapter~\ref{c:threed_sr_b}, both 3-DSM-CYC \cite{Pittel20} and two-dimensional Stable Roommates (SR) \cite{PI94} have been studied from a probabilistic perspective, which are possible starting points. Alternatively, an empirical approach might be informative, as it has been for 3-DSM-CYC \cite{Escamocher2018}. In this direction, it might be useful to formulate the problem of finding a stable matching in a given instance of 3DR-W as an integer programming model.

Various other optimisation problems and measures can be defined in relation to stability and 3DR-W. In Chapter~\ref{c:threed_sr_b} we proposed some alternatives in the setting of 3DR-B, all of which can be defined analogously for 3DR-W.