In this chapter we studied a new formalism of 3DR, involving $\mathscr{B}$-preferences, which we called 3DR-B. We considered in 3DR-B the existence of, and complexity of finding, matchings that are stable. 

Our first result was that a given instance of 3DR-B may not contain a stable matching and that the associated decision problem is $\NP$-complete. We then considered a closely related optimisation problem, which we called 3DR-B-MSM, in which the objective is to construct, in a given instance of 3DR-B, a matching with the maximum number of non-blocking triples. We first devised a $9/4$-approximation algorithm for 3DR-B-MSM based on an existing algorithm for 3PSA-MSM, which is a closely related problem \cite{rosenbaum16}. Improving upon this approximation, we then presented a $3/2$-approximation algorithm based on serial dictatorship, and showed that our analysis is tight asymptotically. Finally, we considered the problem of identifying the smallest instance of 3DR-B that contains no stable matching. We showed that such an instance must have between $9$ and $15$ agents, inclusive.

% As we noted in Chapter~\ref{c:lit_review}, there are a wide variety of other systems of preference representation and solution concepts that could be used to formalise alternative model of 3DR, as well as $\mathscr{B}$, $\mathscr{W}$, and additively separable preferences. 

We now present some open problems specifically involving stability in 3DR-B. More general problems, involving solution concepts other than stability and other models of fixed-size coalition formation, are discussed in Chapter~\ref{c:conclusion}. 

An immediate open problem is to improve the bounds on the the smallest instance of 3DR-B that contains no stable matching. We have shown in Section~\ref{sec:threed_sr_b_structural} that such an instance contains at least $9$ and at most $15$ agents, but the precise number of agents remains open. To fully resolve this open question, it will be necessary to either prove that every some fixed size strictly greater than $6$ contains a stable matching (as in the proof of Theorem~\ref{thm:threed_sr_b_ifnis6thensmexists}), demonstrate that some instance with between $9$ and $12$ agents contains no stable matching (in a similar way to the proof of Theorem~\ref{thm:threed_sr_b_ifnis15thensmdoesnotexist}), or both.

In Chapter~\ref{c:three_dsm_cyc}, we presented an approximation algorithm for a restriction of 3-DSM-CYC-MSM in which the preferences of some agents were derived from a master list. We conjecture that a similar algorithm also exists for a restriction of 3DR-B-MSM in which the preferences of all agents are derived from a master list. In 2020, Bredereck et al.\ \cite{Bre20} considered a similar situation for a multidimensional generalisation of 3GSM \cite{NH91}, and it may be that some of their results or techniques can also be applied to 3DR-B.

A closely related objective is to estimate the probability that a random instance of 3DR-B contains a stable matching. Pittel's \cite{Pittel20} probabilistic analysis of $k$-DSM-CYC and Pittel and Irving's \cite{PI94} analysis of two-dimensional Stable Roommates (SR) are two possible starting points. A hybrid of theoretical and empirical techniques might also be informative, as it was in Escamocher and O'Sullivan's \cite{Escamocher2018} paper, which considered the same question in the setting to 3-DSM-CYC.

As well as investigating the existence of an improved approximation algorithm for 3DR-B-MSM, it might be possible to prove an inapproximability result for this problem. For example, it would be very informative to prove that the approximation ratio of Algorithm~\algorithmfont{serialDictatorship} is tight, by showing that no $(3/2 - \varepsilon)$-approximation algorithm exists for 3DR-B-MSM unless $\P = \NP$. Alternatively, it might be easier to prove the weaker result that 3DR-B-MSM is $\APX$-hard, meaning that there exists some constant factor $\varepsilon$ such that no $(1 + \varepsilon)$-approximation algorithm exists for 3DR-B-MSM, unless $\P = \NP$ \cite{ACGKMP99}. A starting point towards the latter result could be to modify the reduction of Iwama et al.\ \cite{IMO08} from a variant of \emph{Maximum 3D Matching} (Max 3DM) to an optimisation problem defined in a related model of 3DR (which is described in Chapter~\ref{c:lit_review}). Another possibility is to adapt the reduction of Rosenbaum \cite{rosenbaum16} from Max 3DM to 3PSA-MSM.

Various alternative optimisation problems and measures can also be defined that relate to stable matchings and 3DR-B.
One possibility is to construct a sub-matching of maximum cardinality such that no three agents in triples in the sub-matching form a blocking triple in the sub-matching. Rosenbaum \cite{rosenbaum16} refers to the analogous problem for 3PSA as the \emph{3PSA Maximum Stable Sub-matching problem} (3PSA-MSS). 
A second possibility is to consider $\alpha$-stability \cite{ABEOMP09} (discussed in Chapter~\ref{c:lit_review}) in the setting of 3DR-B. For example, for some fixed $\alpha \geq 1$, we could say that a matching $M$ is $\alpha$-stable if for any agent $\alpha_i$ and any triple $t$ where $\alpha_i \in t$ the increase in rank in $P_{\alpha_i}$ from $\mathscr{B}(M)$ to $\mathscr{B}(t)$ is at most $\alpha$. We could also then define an optimisation problem in which the objective is to find an $\alpha$-stable matching for a minimum such $\alpha$.
A third possibility is to define a complementary problem in which the objective is to minimise the number of blocking triples, which is arguably more natural. Similar optimisation problems, in which the objective is to minimise the number of blocking pairs, have been studied in the context of (two-dimensional) Stable Roommates \cite{ABM06}. % Unlike 3DR-B, deciding if a given instance of SR contains a stable matching is solvable in polynomial time, and it is unclear if the techniques relating to SR can be also applied to the minimisation problem of 3DR-B-MSM.

% In terms of parameterised complexity, a natural starting point is a recent work of Bredereck et al.\ \cite{Bre20}, which explores the parameterised complexity of a generalisation of 3GSM in which each agent's preference list, over sets of agents, is derived from a central master list or poset. It may be possible to consider an analogous situation in 3DR-B in which each agent's preference list, of individual agents, is derived from a central master list or poset.